---
title: "Project Proposal"
author: ""
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, message = FALSE, warning = FALSE)
```

## Predicting Term Deposit Subscription in Portuguese Banking Campaigns

***
  
**Research Questions**

1. What client demographic and financial attributes are most predictive of a term deposit subscription?

**Approach:** Use logistic regression and decision tree classifiers to assess the impact of variables like age, job, marital status, and balance on subscription likelihood. 

**Goal: ** Identify client characteristics that strongly influence subscription rates, enabling targeted outreach based on client profiles.

2. Which campaign-related factors (e.g., contact type, duration, timing) are associated with higher subscription rates?

**Approach:** Conduct exploratory data analysis (EDA) and feature importance analysis using random forests to evaluate the effectiveness of campaign attributes (contact type, last contact duration, month of last contact). 

**Goal: ** Discover optimal campaign strategies and timing to increase subscription rates, refining marketing tactics based on campaign characteristics.

3. How do previous campaign interactions (e.g., past contacts, outcomes of previous campaigns) affect the likelihood of a successful subscription?

**Approach:** Analyze past campaign metrics (e.g., previous contacts, days since last contact, previous campaign outcomes) and apply predictive modeling (logistic regression and support vector machines). 

**Goal: ** Assess the impact of prior client interactions on the current campaign’s success, informing strategies for follow-up or re-engagement.


### Project Details

- Cases: Individual clients contacted during the bank’s marketing campaigns.

- Sample Size: 45,211 records in the bank.csv data set (with 17 input features).

### Data Collection Method

- Description: Data is derived from phone-based marketing campaigns conducted by a Portuguese bank between 2008 and 2010, targeting clients for term deposit subscriptions.

- Purpose: Gather information on client and campaign attributes to study factors influencing term deposit subscriptions.

### Type of Study

- Observational Study: This is an observational data set, as it records outcomes without any experimental manipulation or random assignment.

### Data Source

- Source: Obtained from the UCI Machine Learning Repository.

- Citation: Moro, S., Cortez, P., & Rita, P. (2014). A data-driven approach to predict the success of bank telemarketing. Decision Support Systems, 62, 22-31. 

- UCI Repository link: **https://archive.ics.uci.edu/ml/datasets/bank+marketing**

### Response Variable

- Response Variable: y (whether a client subscribed to a term deposit).

- Type: Categorical (Binary: "yes" or "no").

### Explanatory Variables

- Client Attributes: Age (Numerical), Job (Categorical), Marital Status (Categorical), Education (Categorical), Balance (Numerical), Default (Binary), Housing Loan (Binary), Personal Loan (Binary).

- Campaign Details: Contact Type (Categorical), Last Contact Day (Numerical), Last Contact Month (Categorical), Last Contact Duration (Numerical).

- Additional Attributes: Number of Contacts (campaign, Numerical), Days Since Last Contact (pdays, Numerical), Contacts in Previous Campaign (previous, Numerical), Outcome of Last Campaign (poutcome, Categorical).

# Summary Statistics

```{r}
library(tidyverse)
library(readr)
library(dplyr)
```

### Acquiring Data set

```{r}
data = read_delim("https://raw.githubusercontent.com/SrinivasA06/Python/refs/heads/main/bank.csv"
                  , delim = ";")
nrow(data)
glimpse(data)
```

### Numerical Variables

```{r}
# List of numerical columns
numerical_columns <- c("age", "balance", "day", "duration", "campaign", "pdays", "previous")
summary(data[numerical_columns])
```

**Key Takeaways**

- Age: The dataset covers a broad age range, with a slight skew towards middle-aged individuals.
- Balance: The presence of negative balances and a few very high balances suggests varied financial situations among the respondents.
- Campaign Engagement: Most individuals were contacted once or twice during the campaign, with only a few receiving extensive engagement.
- Pdays and Previous Campaign: A large proportion of individuals have -1 (no prior contact) values in both the pdays and previous columns, suggesting limited prior engagement.
- Duration: The call duration varies significantly, with most calls being relatively short, but a few extending much longer.

### Count of Missing Values 

```{r}
# Count missing values per column
missing_counts <- sapply(data, function(x) sum(is.na(x)))
missing_counts
```

### Frequency Table

```{r}
# List of categorical columns
categorical_columns <- setdiff(names(data), numerical_columns)
lapply(data[categorical_columns], table)
```

**Key Takeaways**

- Most respondents are **married** with **secondary education** and are not in loan situations.
- The **cellular contact** method is the most common, and **May** appears to be a peak contact month.





